The Evolution of Phenotypically Invariant Genetic Networks

### 2017-02-07 kalman_walk.tex
added a section on deleterious stuff and a temporary fitness function.

\documentclass[11 pt]{article}
\usepackage{amsmath, amssymb, color, xcolor}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage[backref=page, hidelinks, colorlinks=true, citecolor=blue!60!black!100]{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, shapes}
\usepackage{caption, subcaption}
\usepackage{natbib} % gives us \citet: Author (year) and \citep: (Author; year)
\usepackage{authblk}

\newcommand{\plr}[1]{{\color{blue}\it #1}}
\newcommand{\jss}[1]{{\color{olive}\it #1}}
% \newcommand{\ddt}{\frac{d}{dt}}
\newcommand{\ddt}{\dot}
\newcommand{\ro}{{ro}}
\newcommand{\nro}{{\bar{r}o}}
\newcommand{\rno}{{r\bar{o}}}
\newcommand{\nrno}{{\bar{r}\bar{o}}}
\newcommand{\reachable}{\mathcal{R}}
\newcommand{\unobservable}{\bar{\mathcal{O}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}

@@

>>>>

\documentclass[11 pt]{article}
\usepackage{amsmath, amssymb, color, xcolor}
\usepackage{graphicx, wrapfig, float, caption}
\usepackage{fullpage}
\usepackage[backref=page, hidelinks, colorlinks=true, citecolor=blue!60!black!100]{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, shapes}
\usepackage{caption, subcaption}
\usepackage{natbib} % gives us \citet: Author (year) and \citep: (Author; year)
\usepackage{authblk}

\newcommand{\plr}[1]{{\color{blue}\it #1}}
\newcommand{\jss}[1]{{\color{olive}\it #1}}
% \newcommand{\ddt}{\frac{d}{dt}}
\newcommand{\ddt}{\dot}
\newcommand{\ro}{{ro}}
\newcommand{\nro}{{\bar{r}o}}
\newcommand{\rno}{{r\bar{o}}}
\newcommand{\nrno}{{\bar{r}\bar{o}}}
\newcommand{\reachable}{\mathcal{R}}
\newcommand{\unobservable}{\bar{\mathcal{O}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\pda}{\frac{\partial}{\partial A_{ij}}}

@@

\section{Including Deleterious Mutations and a Fitness Function}

  \begin{minipage}{0.5\textwidth}
  At the moment we focus strictly on the case where the minimal realization is in $2$ dimensions and both of its eigenvalues are real and negative (so a special case of the system from Example \ref{ex:2x2}). \jss{(Note: this is not to say the ideas below will not be generalizable).}

    The system of interest has the form,
  \begin{align*}
    \Sigma := \left\{ A = \begin{bmatrix} A_{11} & A_{12} \\ A_{21} & A_{22} \end{bmatrix}, B = \begin{bmatrix} 1 \\ 1 \end{bmatrix}, C = \begin{bmatrix} 1 & 0 \end{bmatrix} \right\}
  \end{align*}
  \end{minipage}%
  \begin{minipage}{0.5\textwidth}
    \centering
    \includegraphics[width=0.95\textwidth]{figure_2x2mutation01}%
      \captionof{figure}{The transfer function for $A = \left[ \protect\begin{array}{cc} -3 & 1 \\ 0 & -1 \protect\end{array} \right]$ and all eight $\mu(\Sigma, A_{ij})$ mutations, where $\upsilon = \pm 0.5$}\label{fig:mut}%
  \end{minipage}

Define the fitness function to be,

\begin{align}\label{eq:fitness}
  \Phi(\bar{\Sigma}) := \left\{ \begin{array}{ll}   e^{- | H(0) - \tilde{H}(0) | = - \left| \frac{k}{\lambda_{1} \lambda_{2}} - \frac{\tilde{k}}{\tilde{\lambda}_{1} \tilde{\lambda}_{2}} \right|}    &\text{if } \lambda_{1,2} < 0 \text{ and } \in \mathbb{R}, \ \det(A) \neq 0 \\
  0   &\text{if } \lambda_{1,2} \geq 0 \ \text{ or } \ \notin \mathbb{R}, \ \det(A) = 0 \end{array} \right.
\end{align}
and let $\upsilon$ be the magnitude and direction (sign) of a mutation on regulatory interaction strength. The fitness function (\ref{eq:fitness}) may me appropriate if mutations only slightly alter the dynamics of the system, as in Figure \ref{fig:mut}.

Further let $\mu(\Sigma, A_{ij})$ be resulting gene regulatory network following a mutation of matrix element $A_{ij}$ to system $\Sigma$ (which contains matrix $A$).

  \begin{align*}
    \Phi(\mu(\Sigma, A_{ij} \upsilon)) = \exp\left(- \left| \frac{A_{22}-A_{12}}{\lambda_{1} \lambda_{2}} - \frac{k + \upsilon \pda (A_{22}-A_{12})}{\lambda_{1} \lambda_{2} + \text{adj}(A)^{T}_{ij} \upsilon} \right|\right)   .
  \end{align*}
For instance, the fitness of an organism with a change of one of its entries $A_{ij}$ (by $\upsilon$) is listed,
  \begin{align*}
    \Phi(\mu(\Sigma, A_{11})) &= \exp \left(- \left| \frac{A_{22}-A_{12}}{\det(A)} - \frac{A_{22}-A_{12}}{\det(A) + \upsilon A_{22}} \right| \right) \\
    \Phi(\mu(\Sigma, A_{12})) &= \exp \left(- \left| \frac{A_{22}-A_{12}}{\det(A)} - \frac{A_{22}-A_{12} - \upsilon}{\det(A) - \upsilon A_{21}} \right| \right) \\
    \Phi(\mu(\Sigma, A_{21})) &= \exp \left(- \left| \frac{A_{22}-A_{12}}{\det(A)} - \frac{A_{22}-A_{12}}{\det(A) - \upsilon A_{12}} \right| \right) \\
    \Phi(\mu(\Sigma, A_{22})) &= \exp \left(- \left| \frac{A_{22}-A_{12}}{\det(A)} - \frac{A_{22}-A_{12} + \upsilon}{\det(A) + \upsilon A_{11}} \right| \right)
  \end{align*}
